Viral Sequence Sorting Tools Evaluation

### 2022-11-01 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
New visualizations of testing data.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
library(pROC)
```

@@

```{r}
library("stringr")
```

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Bacterial False Positives")
ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Viral False Negatives")
```

@@

@@

@@

>>>>

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
library(pROC)
library("stringr")
```

@@

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Bacterial False Positives")

ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Viral False Negatives")

ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Viral False Negatives")

ggplot(accuracy_scores, aes(x=protist_FP, y=fungal_FP, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Protist FP") +
  ylab("Fungal FP")

ggplot(accuracy_scores, aes(x=recall, y=fungal_FP, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Recall") +
  ylab("Fungal FP")

ggplot(accuracy_scores, aes(x=recall, y=protist_FP, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Recall") +
  ylab("Protist FP")
```

@@

```{r}
TP <- table(viruses$confusion_matrix_high_MCC)[4]
FP <- table(viruses$confusion_matrix_high_MCC)[2]
TN <- table(viruses$confusion_matrix_high_MCC)[3]
FN <- table(viruses$confusion_matrix_high_MCC)[1]

precision <- TP/(TP+FP)
precision
recall <- TP/(TP+FN)
recall
F1 <- 2*precision*recall/(precision+recall)
F1

MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
MCC
```

precision=66%, recall=91%, MCC=74%

precision adjusting size to be equal viral/not viral
```{r}
TP <- table(viruses$confusion_matrix_high_MCC)[4]
FP <- table(viruses$confusion_matrix_high_MCC)[2]*.11
TN <- table(viruses$confusion_matrix_high_MCC)[3]*.11
FN <- table(viruses$confusion_matrix_high_MCC)[1]

precision <- TP/(TP+FP)
precision
recall <- TP/(TP+FN)
recall
F1 <- 2*precision*recall/(precision+recall)
F1

MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
MCC
```

precision=0.94, recall=0.90, F1=0.92, MCC=0.85

@@

```{r}
TP <- table(viruses$confusion_matrix_all)[4]
FP <- table(viruses$confusion_matrix_all)[2]
TN <- table(viruses$confusion_matrix_all)[3]
FN <- table(viruses$confusion_matrix_all)[1]

precision <- TP/(TP+FP)
precision
recall <- TP/(TP+FN)
recall
F1 <- 2*precision*recall/(precision+recall)
F1

MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
MCC
```
precision=62%, recall=92%, MCC=73%

@@

## few tools, high MCC example
```{r}
viruses$keep_score_few_tools <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
```

```{r}
viruses$confusion_matrix_few_tools <- "true negative"
viruses$confusion_matrix_few_tools[viruses$seqtype=="virus" & viruses$keep_score_few_tools<1] <- "false negative"
viruses$confusion_matrix_few_tools[viruses$seqtype=="virus" & viruses$keep_score_few_tools>=1] <- "true positive"
viruses$confusion_matrix_few_tools[viruses$seqtype!="virus" & viruses$keep_score_few_tools>=1] <- "false positive"
```

```{r}
TP <- table(viruses$confusion_matrix_few_tools)[4]
FP <- table(viruses$confusion_matrix_few_tools)[2]
TN <- table(viruses$confusion_matrix_few_tools)[3]
FN <- table(viruses$confusion_matrix_few_tools)[1]

precision <- TP/(TP+FP)
precision
recall <- TP/(TP+FN)
recall
F1 <- 2*precision*recall/(precision+recall)
F1

MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
MCC
```
precision=77%, recall=76%, MCC=74%
